---
title: "MP01_Part02_Fish_Diet_Data"
author: "Jalal Mansoori"
date: "11/8/2020"
output: html_document
---

```{r}
#Loading necessary packages
library(tidyverse)
```

## Exploratory Data Analysis

```{r}
# Loading the fish_diet dataset

df_fishDiet <- read.csv('fish-diet.csv')
str(df_fishDiet)
```
So we have 2 categorical variable fish_in_diet and cancer.

```{r}
#Getting first few values in the dataset
head(df_fishDiet)
```

```{r}
#Count the data with respect to the values in each variable
# For Proportion of fish present in diet
df_fishDiet %>%
  count(fish_in_diet) %>%
  mutate(p = n /sum(n))

```
```{r}
#Plotting the bar plot

ggplot(data= df_fishDiet, aes(x= fish_in_diet)) + 
  geom_bar()
```


```{r}
# For Proportion of Cancer
df_fishDiet %>%
  count(cancer) %>%
  mutate(p = n /sum(n))
```
```{r}
#Plotting the bar plot

ggplot(data= df_fishDiet, aes(x= cancer)) + 
  geom_bar()
```

## Estimating the True Population proportion for swedish men that have prostate cancer based on the given dataset (considering any amount of fish diet)


We know that the sample proportion of Swedish men having cancer is 7.42%.

```{r}
#Calculating 95% confidence interval for true population proportion of Swedish men having cancer (considering any amount of fish present in their diet)


phat<- 0.0742
standard_error <- sqrt(phat*(1 - phat)/length(df_fishDiet$cancer)) 

phat + c(-1, 1) * qnorm(0.975)*standard_error
```
We are 95% Confident that the true population proportion of Swedish men having cancer is in between 6.8% and 8.1%




